Location and Scale-Invariant Power Transformations for Transforming Data to Normality

### 2025-10-29 manuscript_latex/manuscript.tex
Update appendix d, part 2.

The sequences drawn resulted from the permutations of the number of
elements of each sequence (\(n \in \{30, 100, 500\}\)) and
transformation parameter for the inverse transformation
\(\lambda \in \{0.1, 1.0, 1.9\}\). Each sequence then underwent power
transformation using Box-Cox and Yeo-Johnson transformations. For
Box-Cox transformations, each sequence was shifted prior to inverse
transformation so that every element after inverse transformation would
be strictly positive.

@@

\begin{table}
\caption{Residual errors for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-appendix-residuals-yj}
\small{
\begin{tabular}{l | l r r r r r r r r r}

@@

\bottomrule
\end{tabular}
}
\end{table}

\begin{table}
\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-appendix-lambda-yj}
\small{
\begin{tabular}{l | l r r r r r r r r r}

@@

conventional                          & & 0.6 & 1.2 & 1.5 &  0.0 & 0.9 & 1.6 &  0.1 & 1.0 & 1.9 \\
conventional (z-score norm.)          & & 0.5 & 1.2 & 1.5 & -0.2 & 0.9 & 1.7 & -0.2 & 1.0 & 2.2 \\
conventional (robust scaling)         & & 0.5 & 1.3 & 1.6 & -0.3 & 0.9 & 1.7 & -0.2 & 1.0 & 2.1 \\
Raymaekers-Rousseeuw                  & & 1.0 & 1.2 & 1.5 &  0.0 & 0.7 & 1.6 &  0.1 & 1.0 & 1.9 \\
Raymaekers-Rousseeuw (z-score norm.)  & & 0.7 & 1.2 & 1.5 & -0.2 & 0.7 & 1.7 & -0.2 & 1.0 & 2.2 \\
Raymaekers-Rousseeuw (robust scaling) & & 0.6 & 1.3 & 1.6 & -0.3 & 0.6 & 1.7 & -0.2 & 1.0 & 2.1 \\
invariant                             & & 0.4 & 1.4 & 1.7 &  0.1 & 0.9 & 1.8 &  0.1 & 1.1 & 3.0 \\
robust invariant                      & & 0.5 & 1.1 & 1.5 & -0.1 & 0.7 & 1.7 &  0.1 & 1.1 & 1.9 \\
\bottomrule
\end{tabular}
}
\end{table}

\begin{table}
\caption{
p-values of empirical central normality tests for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-appendix-p-value-yj}
\small{
\begin{tabular}{p{3.5cm} | l r r r r r r r r r}

@@

none                                  & & 0.981 & 0.816 & 0.576 & $<0.001$ & 0.099 & 0.011 & $<0.001$ & 0.993 & $<0.001$ \\
conventional                          & & 0.979 & 0.614 & 0.617 &   0.722  & 0.221 & 0.782 &   0.946  & 0.997 &   0.637 \\
conventional (z-score norm.)          & & 0.970 & 0.594 & 0.645 &   0.251  & 0.224 & 0.895 &   0.529  & 0.998 &   0.828 \\
conventional (robust scaling)         & & 0.974 & 0.593 & 0.640 &   0.538  & 0.207 & 0.837 &   0.933  & 0.998 &   0.810 \\
Raymaekers-Rousseeuw                  & & 0.983 & 0.614 & 0.617 &   0.722  & 0.622 & 0.782 &   0.927  & 0.985 &   0.831 \\
Raymaekers-Rousseeuw (z-score norm.)  & & 0.985 & 0.594 & 0.645 &   0.251  & 0.623 & 0.895 &   0.564  & 0.984 &   0.807 \\
Raymaekers-Rousseeuw (robust scaling) & & 0.987 & 0.593 & 0.640 &   0.538  & 0.599 & 0.837 &   0.935  & 0.984 &   0.875 \\
invariant                             & & 0.987 & 0.487 & 0.573 &   0.778  & 0.241 & 0.683 &   0.963  & 0.997 &   0.865 \\
robust invariant                      & & 0.992 & 0.753 & 0.667 &   0.909  & 0.739 & 0.628 &   0.968  & 0.994 &   0.680 \\

\bottomrule
\end{tabular}
}
\end{table}

\begin{table}
\caption{Residual errors for features from simulated clean data without outliers after Box-Cox transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-appendix-residuals-bc}
\small{
\begin{tabular}{l | l r r r r r r r r r}

@@

\bottomrule
\end{tabular}
}
\end{table}

\begin{table}
\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Box-Cox transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-appendix-lambda-bc}
\small{
\begin{tabular}{l | l r r r r r r r r r}

@@

\bottomrule
\end{tabular}
}
\end{table}

\begin{table}
\caption{
p-values of empirical central normality tests for features from simulated clean data without outliers after Box-Cox transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-appendix-p-value-bc}
\small{
\begin{tabular}{p{3.5cm} | l r r r r r r r r r}

@@

none                                  & & 0.940 & 0.816 & 0.498 & $<0.001$ & 0.098 & 0.714 & $<0.001$ & 0.993 & 0.928 \\
conventional                          & & 0.948 & 0.695 & 0.603 &   0.798  & 0.163 & 0.791 &   0.931  & 0.997 & 0.658 \\
conventional (z-score norm.)          & & 0.947 & 0.777 & 0.624 &   0.611  & 0.182 & 0.788 &   0.969  & 0.991 & 0.822 \\
conventional (robust scaling)         & & 0.947 & 0.777 & 0.624 &   0.611  & 0.182 & 0.788 &   0.969  & 0.991 & 0.822 \\
Raymaekers-Rousseeuw                  & & 0.948 & 0.695 & 0.603 &   0.798  & 0.768 & 0.806 &   0.816  & 0.962 & 0.782 \\
Raymaekers-Rousseeuw (z-score norm.)  & & 0.947 & 0.777 & 0.753 &   0.913  & 0.255 & 0.784 &   0.775  & 0.866 & 0.921 \\
Raymaekers-Rousseeuw (robust scaling) & & 0.947 & 0.777 & 0.753 &   0.913  & 0.255 & 0.784 &   0.807  & 0.866 & 0.921 \\
invariant                             & & 0.947 & 0.661 & 0.624 &   0.611  & 0.182 & 0.789 &   0.956  & 0.997 & 0.664 \\
robust invariant                      & & 0.969 & 0.824 & 0.750 &   0.972  & 0.779 & 0.775 &   0.956  & 0.934 & 0.840 \\

\bottomrule
\end{tabular}
}
\end{table}

>>>>

The sequences drawn resulted from the permutations of the number of
elements of each sequence (\(n \in \{30, 100, 500\}\)) and
transformation parameter for the inverse transformation
\(\lambda \in \{0.1, 1.0, 1.9\}\). Each sequence then underwent power
transformation using Box-Cox and Yeo-Johnson transformations. For
Box-Cox transformations, each sequence was shifted prior to inverse
transformation ensuring all elements were strictly positive after
inverse transformation.

@@

\begin{table}
\begin{center}
\caption{Residual errors for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-appendix-residuals-yj}
\small{
\begin{tabular}{l | l r r r r r r r r r}

@@

\bottomrule
\end{tabular}
}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-appendix-lambda-yj}
\small{
\begin{tabular}{l | l r r r r r r r r r}

@@

conventional                          & & 0.6 & 1.2 & 1.5 &  0.0 & 0.9 & 1.6 &  0.1 & 1.0 & 1.9 \\
conventional (z-score norm.)          & & 0.5 & 1.2 & 1.5 & -0.2 & 0.9 & 1.7 & -0.2 & 1.0 & 2.2 \\
conventional (robust scaling)         & & 0.5 & 1.3 & 1.6 & -0.3 & 0.9 & 1.7 & -0.2 & 1.0 & 2.1 \\
Raymaekers-Rousseeuw                  & & 1.0 & 1.2 & 1.5 &  0.0 & 0.7 & 1.6 &  0.1 & 1.0 & 1.9 \\
Raymaekers-Rousseeuw (z-score norm.)  & & 0.7 & 1.2 & 1.5 & -0.2 & 0.7 & 1.7 & -0.2 & 1.0 & 2.2 \\
Raymaekers-Rousseeuw (robust scaling) & & 0.6 & 1.3 & 1.6 & -0.3 & 0.6 & 1.7 & -0.2 & 1.0 & 2.1 \\
invariant                             & & 0.4 & 1.4 & 1.7 &  0.1 & 0.9 & 1.8 &  0.1 & 1.1 & 3.0 \\
robust invariant                      & & 0.5 & 1.1 & 1.5 & -0.1 & 0.7 & 1.7 &  0.1 & 1.1 & 1.9 \\
\bottomrule
\end{tabular}
}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{
p-values of empirical central normality tests ($\kappa = 0.80$) for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-appendix-p-value-yj}
\small{
\begin{tabular}{p{3.5cm} | l r r r r r r r r r}

@@

none                                  & & 0.968 & 0.742 & 0.450 & $<0.001$ & 0.058 & 0.004 & $<0.001$ & 0.963 & $<0.001$ \\
conventional                          & & 0.966 & 0.494 & 0.498 &   0.632  & 0.145 & 0.707 &   0.828  & 0.986 &   0.358 \\
conventional (z-score norm.)          & & 0.955 & 0.471 & 0.529 &   0.169  & 0.147 & 0.853 &   0.260  & 0.986 &   0.603 \\
conventional (robust scaling)         & & 0.957 & 0.470 & 0.523 &   0.674  & 0.146 & 0.775 &   0.828  & 0.986 &   0.347 \\
Raymaekers-Rousseeuw                  & & 0.972 & 0.494 & 0.498 &   0.632  & 0.518 & 0.707 &   0.785  & 0.939 &   0.609 \\
Raymaekers-Rousseeuw (z-score norm.)  & & 0.974 & 0.471 & 0.529 &   0.169  & 0.519 & 0.853 &   0.289  & 0.935 &   0.568 \\
Raymaekers-Rousseeuw (robust scaling) & & 0.968 & 0.470 & 0.523 &   0.674  & 0.536 & 0.775 &   0.762  & 0.935 &   0.609 \\
invariant                             & & 0.978 & 0.353 & 0.448 &   0.701  & 0.161 & 0.588 &   0.872  & 0.984 &   0.667 \\
robust invariant                      & & 0.985 & 0.660 & 0.556 &   0.871  & 0.652 & 0.525 &   0.888  & 0.968 &   0.411 \\

\bottomrule
\end{tabular}
}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{Residual errors for features from simulated clean data without outliers after Box-Cox transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-appendix-residuals-bc}
\small{
\begin{tabular}{l | l r r r r r r r r r}

@@

\bottomrule
\end{tabular}
}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Box-Cox transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-appendix-lambda-bc}
\small{
\begin{tabular}{l | l r r r r r r r r r}

@@

\bottomrule
\end{tabular}
}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{
p-values of empirical central normality tests ($\kappa = 0.80$) for features from simulated clean data without outliers after Box-Cox transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-appendix-p-value-bc}
\small{
\begin{tabular}{p{3.5cm} | l r r r r r r r r r}

@@

\midrule
none                                  & & 0.911 & 0.742 & 0.365 & $<0.001$ & 0.057 & 0.623 & $<0.001$ & 0.962 & 0.789 \\
conventional                          & & 0.922 & 0.590 & 0.482 &   0.726  & 0.100 & 0.718 &   0.794  & 0.984 & 0.380 \\
conventional (z-score norm.)          & & 0.920 & 0.690 & 0.506 &   0.505  & 0.114 & 0.714 &   0.888  & 0.957 & 0.594 \\
conventional (robust scaling)         & & 0.920 & 0.690 & 0.506 &   0.504  & 0.114 & 0.714 &   0.887  & 0.957 & 0.594 \\
Raymaekers-Rousseeuw                  & & 0.922 & 0.590 & 0.482 &   0.726  & 0.689 & 0.736 &   0.582  & 0.869 & 0.531 \\
Raymaekers-Rousseeuw (z-score norm.)  & & 0.920 & 0.690 & 0.660 &   0.876  & 0.172 & 0.710 &   0.521  & 0.670 & 0.774 \\
Raymaekers-Rousseeuw (robust scaling) & & 0.920 & 0.690 & 0.660 &   0.876  & 0.172 & 0.710 &   0.568  & 0.670 & 0.774 \\
invariant                             & & 0.921 & 0.549 & 0.506 &   0.505  & 0.114 & 0.715 &   0.853  & 0.986 & 0.387 \\
robust invariant                      & & 0.954 & 0.752 & 0.656 &   0.956  & 0.699 & 0.698 &   0.852  & 0.801 & 0.623 \\

\bottomrule
\end{tabular}
}
\end{center}
\end{table}
